---
title: "Regression with PC's as dependent variable"
author: '510646291'
date: "2024-10-10"
output: html_document
---

## Load data and Flynn's work

```{r}
library(tidyverse)
library(knitr)
library(here)
library(MASS)
i_am("wide_adults_food_servings_foods_grouped.RData")
load(here("wide_adults_food_servings_foods_grouped.RData"))
load(here("PCA_results.RData"))
load(here("..","data","tech_data.RData"))
```

## Data management

We combined PCA results with clinical outcomes.

```{r}
plot(pca_result,type = "l")
pc.df <- as.data.frame(pca_result$x) %>% dplyr::select(paste0("PC",1:5))
combined.df1 <- cbind(pc.df,wide_adults_food_servings_foods_grouped) %>% dplyr::select(ABSPID,paste0("PC",1:5))
combined.df2 <- left_join(combined.df1,tech_biom,by="ABSPID")
```
From this graph, it seems that analyzing the first five PCS is enough."combined.df2" will be the data to analyse.

## Data cleaning

Data cleaning includes the following steps:

- set our of range values to missing
- drop variables with missingness greater than 30%
- only keep complete cases for next regression step

Clean data contain 2705 samples with 5 PCs and 33 clinical outcomes.

```{r}
clean.df1 <- combined.df2 %>% mutate(
  BMISC = na_if(as.numeric(BMISC), 0) %>% na_if(98) %>% na_if(99),
  AGEC = na_if(as.numeric(AGEC), 99),
  SMSBC = na_if(as.character(SMSBC), "0"),
  FEMLSBC = na_if(as.character(FEMLSBC), "6") %>% na_if("9"),
  PHDKGWBC = na_if(as.numeric(PHDKGWBC), 997) %>% na_if(998) %>% na_if(999),
  PHDCMHBC = na_if(as.numeric(PHDCMHBC), 998) %>% na_if(999),
  EXLWTBC = na_if(as.numeric(EXLWTBC), 9999) %>% na_if(0),
  EXLWMBC = na_if(as.numeric(EXLWMBC), 9999) %>% na_if(0),
  EXLWVBC = na_if(as.numeric(EXLWVBC), 9999) %>% na_if(0),
  PHDCMWBC = na_if(as.numeric(PHDCMWBC), 998) %>% na_if(999),
  SF2SA1QN = na_if(as.character(SF2SA1QN), "99"),
  INCDEC  = na_if(as.character(INCDEC), "0") %>% na_if("99") %>% na_if("98"),
  DIETRDI = na_if(as.character(DIETRDI), "0")
)

missing_pct <- clean.df1 %>% 
  summarise(across(everything(), ~mean(is.na(.)))) 
clean.df2 <- clean.df1 %>% 
  dplyr::select(where(~mean(is.na(.)) <= 0.3))
clean.df3 <- clean.df2[complete.cases(clean.df2), ]

clean.df4 <- clean.df3 %>%
  mutate(across(ends_with("_MISS"), ~ {
    factor_var <- as.factor(.)
    if (nlevels(factor_var) <= 1) {
      return(NULL)
    } else {
      return(factor_var)
    }
  }))
dim(clean.df4)
```

## Variable selection and regression summary

We used a stepwise method with both directions to select variables that are associated with PCs

```{r,include=F}
lm.pc1 <- lm(PC1 ~ .,data=clean.df4[,c(2,7:39)])
lm.pc2 <- lm(PC2 ~ .,data=clean.df4[,c(3,7:39)])
lm.pc3 <- lm(PC3 ~ .,data=clean.df4[,c(4,7:39)])
lm.pc4 <- lm(PC4 ~ .,data=clean.df4[,c(5,7:39)])
lm.pc5 <- lm(PC5 ~ .,data=clean.df4[,c(6,7:39)])

step1 <- step(lm.pc1,direction = "both")
step2 <- step(lm.pc2,direction = "both")
step3 <- step(lm.pc3,direction = "both")
step4 <- step(lm.pc4,direction = "both")
step5 <- step(lm.pc5,direction = "both")
```
Variables associated with PCs are listed below:

- PC1: 0.021938AGEC, -0.780173COBBC3 + 0.006192PHDKGWBC + 0.009589PHDCMHBC + -0.00396EXLWVBC +  -0.350385SF2SA1QN4 + -0.364374SF2SA1QN5+ 0.536230BDYMSQ042 +  0.431977BDYMSQ045+ 0.853609DIETQ56 +0.212464DIETQ53 + 0.262575DIETQ54 +  -0.191842 DIETQ82+ -0.314453DIETQ83+ -0.588822DIETQ85 + 0.283527DIETRDI2 + -0.435377SEX2 + -0.248077SMKDAILY3 +  0.002282SYSTOL + 0.100796BIORESPC2 -3.671429

Explanation: AGEC - Age of person 
             COBBC3 - Country of birth (No English speaking as main country)
             PHDKGWBC - Measured weight (kg)
             PHDCMHBC - Measured height (cm)
             EXLWVBC - Total mins undertaken physical activity in last week
             SF2SA1QN4 -  Decile 4th Index of Relative Socio-Economic Disadvantage
             SF2SA1QN5 - Decile 5th Index of Relative Socio-Economic Disadvantage
             BDYMSQ042 - Currently on a diet for health reasons
             BDYMSQ045 - Not currently on a diet
             DIETQ56 - Usual daily 6 or more serves serves of vegetables
             DIETQ53 - Usual daily 3 serves serves of vegetables
             DIETQ54 - Usual daily 4 serves serves of vegetables
             DIETQ82 - Usual daily 2 serves of fruit
             DIETQ83 - Usual daily 3 serves of fruit
             DIETQ85 - Usual daily 5 serves of fruit
             DIETRDI2 - vegetable and fruit consumption did no met recommended dietary guidelines
             SEX2 -  Female
             SMKDAILY3 - Past daily smoker (ex-smoker)
             SYSTOL - Systolic blood pressure
             BIORESPC2 - Biomedical participant
```{r}
summary(step1)
```

- PC2: -0.0205580BMISC +  0.0061739AGEC + 0.0824003SMSBC2 + 0.3713002COBBC3 + 0.0017931EXLWTBC + 0.0053164EXLWVBC + 
    + -0.2089820BDYMSQ045 + 0.3406298DIETQ53+ 0.5353456DIETQ54+  0.9634038DIETQ55+ 1.5891992DIETQ56 + 0.4033966DIETQ82+ 0.4256818DIETQ83+ 0.6709792DIETQ84+ 0.7098542DIETQ85+ 1.1540151DIETQ86+ 0.3266059DIETRDI2 + 0.1425943SEX2 + 0.7894439SMKDAILY2 +  0.2904375SMKDAILY3+ 0.3630767SMKDAILY4+ 0.3094551SMKDAILY5 + 1.0603718
    
Explanation: BMISC - Body mass index (BMI)
             AGEC - Age of person
             SMSBC2 - Married in a de facto marriage
             COBBC3 - Country of birth (No English speaking as main country)
             EXLWTBC - Total mins undertaken physical activity in last week
             EXLWVBC - Total mins undertaken vigorous physical activity in last week
             BDYMSQ045 - Not currently on a diet
             DIETQ53 - Usual daily 3 serves of vegetables
             DIETQ54 - Usual daily 4 serves of vegetables
             DIETQ55 - Usual daily 5 serves of vegetables
             DIETQ56 - Usual daily 6 serves of vegetables
             DIETQ82 - Usual daily 2 serves of fruit
             DIETQ83 - Usual daily 3 serves of fruit
             DIETQ84 - Usual daily 4 serves of fruit
             DIETQ85 - Usual daily 5 serves of fruit
             DIETQ86 - Usual daily 6 or more serves of fruit
             DIETRDI2 - vegetable and fruit consumption did no met recommended dietary guidelines 
             SEX2 -  Female
             SMKDAILY2 - Past daily smoker (still smokes)
             SMKDAILY3 - Past daily smoker (ex-smoker)
             SMKDAILY4 - Never smoked daily
             SMKDAILY5 - Never smoked

```{r}
summary(step2)
```
- PC3: BMISC + AGEC + SMSBC + PHDKGWBC + EXLWTBC + 
    EXLWVBC + PHDCMWBC + INCDEC + DIABBC + DIETQ5 + DIETQ8 + 
    DIETRDI + SLPTIME + SMKDAILY + BIORESPC
```{r}
summary(step3)
```

```{r}
summary(step4)
```
- PC4: SMSBC + COBBC + PHDCMHBC + EXLWMBC + EXLWVBC + 
    PHDCMWBC + INCDEC + DIABBC + HCHOLBC + DIETQ8 + SEX + BIORESPC

```{r}
summary(step5)
```

- PC5:  BMISC + AGEC + COBBC + PHDCMHBC + EXLWTBC + 
    PHDCMWBC + HSUGBC + ADTOTSE + BDYMSQ04 + DIETQ5 + SEX