# Code for generating principal components, based on food data.

> Loading in John's data:


```{r, message=FALSE}
library(tidyverse)
load("../data/tech_data.Rdata")
```


## Creating dataset

Currently:
- Population is all adults over 45 who responded to the survey on BOTH days.
    - Motivation for this is that metabolism generally decreases with age... probably good to consult the nutrition students, though.
- Food codes have been replaced with their actual labels in the data dictionary.

```{r}
library(stringr)

fooditems_dict_major = read.csv("../data/nutmstatDataItems2019_CSV.csv") |>
  filter(str_length(Category.Code) == 3) |>
  dplyr::rename("THRDIG" = "Category.Code") |>
  dplyr::rename("Category.Label.Major" = "Category.Label")

fooditems_dict_minor = read.csv("../data/nutmstatDataItems2019_CSV.csv") |>
  filter(str_length(Category.Code) == 5) |>
  dplyr::rename("FIVDIG" = "Category.Code") |>
  dplyr::rename("Category.Label.Minor" = "Category.Label")

both_days_adults_food <- tech_food |>
  group_by(ABSPID) |>
  left_join(y = tech_biom |> select(ABSPID, AGEC), by = "ABSPID") |>
  filter(n_distinct(DAYNUM) == 2, AGEC >= 45) |>
  ungroup() |>
  left_join(fooditems_dict_major, by = "THRDIG") |>
  left_join(fooditems_dict_minor, by = "FIVDIG") |>
  select(ABSPID, Category.Label.Major, Category.Label.Minor, GRAMWGT)

head(both_days_adults_food, 10)
```


### (NEW) Splitting necessary major groups by minor groups
```{r}
complex_group_splits <- both_days_adults_food |>
  mutate(Category.Label.Major = 
           ifelse(Category.Label.Major == "Potatoes",
             ifelse(Category.Label.Minor == "Potato products", "Fried Potatoes and chips", "Potatoes and potato dishes"), 
           Category.Label.Major)) |>
  mutate(Category.Label.Major = 
           ifelse(Category.Label.Major == "Mixed dishes where cereal is the major ingredient",
                  ## then
                  case_when(
                    str_detect(Category.Label.Minor, regex("pizza", ignore_case = TRUE)) ~ "Pizza",
                    str_detect(Category.Label.Minor, regex("burgers", ignore_case = TRUE)) ~ "Burgers",
                    str_detect(Category.Label.Minor, regex("sushi", ignore_case = TRUE)) ~ "Sushi",
                    str_detect(Category.Label.Minor, regex("saturated fat >5 g/100 g", ignore_case = TRUE)) ~ "High-fat cereal-based takeaway-style food",
                    # LOW FAT TAKEAWAY: Note: R does not recognise ≤ sign. All higher fat cases captured above.
                    str_detect(Category.Label.Minor, regex("5 g/100 g", ignore_case = TRUE)) | Category.Label.Minor == "Other savoury grain dishes" ~ "Low-fat cereal-based takeaway-style food",
                    .default = "Gnocchi and steamed buns"
                    ),
                  Category.Label.Major
          )
         ) |>
  select(-Category.Label.Minor) |>
  dplyr::rename("Category.Label" = "Category.Label.Major")
```


### Creating a "servings" measure based on grams
Looking at the food data, we can create define "1 serving" as the median grams consumed of a particular item.

From here, we just compute servings as (using tea as an example):

$$\text{Servings(Tea)} = \frac{\text{Recorded tea weight (g)}}{\text{Median tea weight (g)}}$$

By comparing grams to "servings", we can directly compare different types of food consumed, without having to worry about how different foods are usually consumed in different weights.

```{r}
both_days_adults_food_servings <- complex_group_splits |>
  group_by(Category.Label) |> 
  mutate(median_grams = median(GRAMWGT)) |>
  mutate(Servings = ifelse(median_grams != 0, GRAMWGT / median_grams, 0)) # sets to 0 if median grams is 0... avoids any divide by 0 errors.

both_days_adults_food_servings$Servings <- format(both_days_adults_food_servings$Servings, scientific = FALSE)
head(both_days_adults_food_servings, 10)
```

#### Summarising data such that we have one row per individual

Each row is summed servings of food that person consumed over two days.

```{r}
wide_adults_food_servings = both_days_adults_food_servings |>
  select(-GRAMWGT, -median_grams) |>
  mutate(Servings = as.numeric(Servings)) |>
  group_by(ABSPID, Category.Label) |>
  summarise(Total_Servings = sum(Servings, na.rm = TRUE), .groups = 'drop') |>
  pivot_wider(names_from = Category.Label, values_from = Total_Servings, values_fill = 0)

head(wide_adults_food_servings, 10)
```

#### Grouping food items together with substantial overlap

Merging together overlapping food items (columns) aids w factor analysis.

**Note:** Ingredients with minimal nutritional content (e.g. baking soda, herbs) and/or low count will be removed.

```{r}
wide_adults_food_servings_foods_grouped <- wide_adults_food_servings |>
  # Removing variables with minimal nutritional content and/or rare groups, which don't fit in with anything else.
  select(-c(
    `Herbs, spices, seasonings and stock cubes`, # this also includes added salt..? Should be careful.
    `Essences`,
    `Stuffings`,
    `Chemical raising agents and cooking ingredients`,
    `Infant foods`,
    `Infant cereal products`,
    `Cheese substitute`,
    `Soy-based ice confection`,
    `Dry soup mix`,
    `Soy-based yoghurts`,
    `Formula dietary foods`
    )) |>
  # Soups
  mutate(Soups = rowSums(across(matches("Soup", ignore.case = TRUE)), na.rm = TRUE)) |>
  select(-matches("Soup"), Soups) |>
  # Fish
  mutate(`Fish and Fish Products` = rowSums(across(c(
    `Other sea and freshwater foods`, 
    `Crustacea and molluscs (excluding commercially sterile)`, 
    `Fish and seafood products (homemade and takeaway)`,
    `Mixed dishes with fish or seafood as the major component`,
    `Packed (commercially sterile) fish and seafood`,
    `Fin fish (excluding commercially sterile)`
    )), na.rm = TRUE)) |>
  select(-c(
    `Other sea and freshwater foods`, 
    `Crustacea and molluscs (excluding commercially sterile)`, 
    `Fish and seafood products (homemade and takeaway)`,
    `Mixed dishes with fish or seafood as the major component`,
    `Packed (commercially sterile) fish and seafood`,
    `Fin fish (excluding commercially sterile)`)) |>
  # Fruit
  mutate(`Fresh Fruit and Fruit-based Dishes` = rowSums(across(c(
    `Berry fruit`, 
    `Stone fruit`,
    `Citrus fruit`,
    `Other fruit`,
    `Pome fruit`,
    `Tropical and subtropical fruit`,
    `Mixtures of two or more groups of fruit`
    )))) |>
  select(-`Berry fruit`, -`Stone fruit`, -`Citrus fruit`, -`Other fruit`, -`Pome fruit`, -`Tropical and subtropical fruit`, -`Mixtures of two or more groups of fruit`) |>
  # Fruit and muesli bars
  mutate(`Fruit and Muesli-style bars` = rowSums(across(c(
    `Fruit, nut and seed-bars`,
    `Muesli or cereal style bars`)))) |>
  select(-c(`Fruit, nut and seed-bars`, 
            `Muesli or cereal style bars`)) |>
  # Cordials and soft drink
  mutate(`Soft drinks, Cordials, Sports drinks, and Milkshakes` = rowSums(across(c(
    `Soft drinks, and flavoured mineral waters`, 
    `Cordials`, 
    `Flavoured milks and milkshakes`, 
    `Electrolyte, energy and fortified drinks`)))) |>
  select(-c(`Soft drinks, and flavoured mineral waters`, 
            `Cordials`, 
            `Flavoured milks and milkshakes`, 
            `Electrolyte, energy and fortified drinks`)) |>
  # Desserts
  mutate(`Milk-based and other dessert items` = rowSums(across(c(
    `Other dishes where milk or a milk product is the major component`, 
    `Sweet biscuits`, 
    `Frozen milk products`, 
    `Dishes and products other than confectionery where sugar is the major component`, 
    `Custards`,
    `Batter-based products`,
    `Mixed dishes where fruit is the major component`,
    `Cakes, muffins, scones, cake-type desserts`)))) |>
  select(-c(`Other dishes where milk or a milk product is the major component`, 
            `Sweet biscuits`, 
            `Frozen milk products`, 
            `Dishes and products other than confectionery where sugar is the major component`,
            `Mixed dishes where fruit is the major component`,
            `Custards`, `Batter-based products`, `Cakes, muffins, scones, cake-type desserts`)) |>
  # Chocolate and lollies
  mutate(`Chocolate and lollies` = rowSums(across(c(
    `Chocolate and chocolate-based confectionery`, 
    `Other confectionery`)))) |>
  select(-c(`Chocolate and chocolate-based confectionery`, 
            `Other confectionery`)) |>
  # Butters, margarines, and oils
  mutate(`Butters, margarines, and oils` = rowSums(across(c(
    `Butters`,
    `Margarine and table spreads`, 
    `Plant oils`,
    `Unspecified fats`,
    `Dairy blends`,
    `Other fats`)))) |>
  select(-c(`Butters`, `Margarine and table spreads`, `Plant oils`, `Unspecified fats`, `Dairy blends`, `Other fats`)) |>
  # Sauces
  mutate(`Sauces` = rowSums(across(c(`Gravies and savoury sauces`, `Salad dressings`)))) |>
  select(-c(`Gravies and savoury sauces`, `Salad dressings`)) |>
  # Coffee and Tea
  mutate(`Coffee and Tea` = rowSums(across(c(`Coffee and coffee substitutes`, `Tea`)))) |>
  select(-c(`Coffee and coffee substitutes`, `Tea`)) |>
  # Starchy snacks
  mutate(`Starchy snacks` = rowSums(across(c(
    `Extruded or reformed snacks`,
    `Other snacks`,
    `Corn snacks`,
    `Potato snacks`
    )))) |>
  select(-c(
    `Extruded or reformed snacks`,
    `Other snacks`,
    `Corn snacks`,
    `Potato snacks`)) |>
  # Alcohol
  mutate(`Alcohol` = rowSums(across(c(
    `Wines`,
    `Beers`,
    `Spirits`,
    `Cider and perry`,
    `Other alcoholic beverages`
  )))) |>
  select(-c(
    `Wines`,
    `Beers`,
    `Spirits`,
    `Cider and perry`,
    `Other alcoholic beverages`
  )) |>
  # Legumes
  mutate(`Legumes and pulses` = rowSums(across(c(
    `Peas and beans`,
    `Mature legumes and pulses`,
    `Mature legume and pulse products and dishes`
  )))) |>
  select(-c(
    `Peas and beans`,
    `Mature legumes and pulses`,
    `Mature legume and pulse products and dishes`
  )) |>
  # Breakfast cereal and porridge
  mutate(`Breakfast cereal and porridge` = rowSums(across(c(
    `Breakfast cereals, ready to eat`,
    `Breakfast cereals, hot porridge style`
  )))) |>
  select(-c(
    `Breakfast cereals, ready to eat`,
    `Breakfast cereals, hot porridge style`)) |>
  # Cheese, yogurts
  mutate(`Cheese, yogurt, cream` = rowSums(across(c(
    `Cheese`,
    `Yoghurt`,
    `Cream`
  )))) |>
  select(-c(
    `Cheese`,
    `Yoghurt`,
    `Cream`
  )) |>
  # Processed meat
  mutate(`Processed meats` = rowSums(across(c(
    `Processed meat`,
    `Sausages, frankfurts and saveloys`
  )))) |>
  select(-c(
    `Processed meat`,
    `Sausages, frankfurts and saveloys`
  )) |>
  # Red meat, unprocessed
  mutate(`Red meat, unprocessed` = rowSums(across(c(
    `Beef, sheep and pork, unprocessed`,
    `Mammalian game meats`,
  )))) |>
  select(-c(
    `Beef, sheep and pork, unprocessed`,
    `Mammalian game meats`,
  )) |>
  # Eggs and egg dishes
  mutate(`Eggs and egg dishes` = rowSums(across(c(
    `Eggs`,
    `Dishes where egg is the major ingredient`,
  )))) |>
  select(-c(
    `Eggs`,
    `Dishes where egg is the major ingredient`,
  ))

# This is actually quite annoying... `mixed dishes where cereal is main component` includes sandwiches, sushi, pizza, etc... 

## Testing
colnames(wide_adults_food_servings_foods_grouped)
```

> Results of variable merging (so far).

```{r}
colnames(wide_adults_food_servings_foods_grouped)
```


## PCA

```{r}
pca_result <- prcomp(wide_adults_food_servings_foods_grouped |> select(-ABSPID), scale = TRUE)

summary(pca_result)

# This is what is used in regression.
pca_scores <- pca_result$x

# loadings (contributions of each variable to the components)
pca_loadings <- pca_result$rotation

# Plot the variance explained by each principal component
plot(pca_result, type = "l")
```

There's approx. 4 PC's with eigenvalue > 1.5.

#### Saving PCA and food serving data

```{r}
save(pca_result, file = "PCA_results.RData")
save(wide_adults_food_servings_foods_grouped,file = "wide_adults_food_servings_foods_grouped.RData")
```



